TTS OD by Purpose by Start Time/Day of Week

### 2023-11-09 data-raw/TTS16-data-inputs/OD-GGH4-by-Time-Day-Mode-Purpose/Extract-OD-by-Purpose-Mode-Time-Day-from-Raw-Files.Rmd
Adding preprocessing files for newest table

Loop through the tables and check the filters recorded in each:
```{r}
# Parameters for testing, commented out
d <- 1
p <- 2
m <- 6

@@

for(p in 1:length(purpose)){
  # Loop through modes
  for(m in 1:length(mode)){
    # Loop through weekdays
    for(d in 1:length(weekday)){
      # Read the files in turn
      raw_data <- read_delim(file = glue("TTS16-data-inputs/OD-GGH4-by-Time-Day-Mode-Purpose/ttsCross-OD-GGH4-{purpose[p]}-{weekday[d]}-{mode[m]}.txt"), 
                             delim = "\t", 
                             col_names = FALSE,
                             show_col_types = FALSE)

@@

# Loop through purposes
for(p in 1:length(purpose)){
  # Loop through modes
  for(m in 1:length(mode)){
    # Loop through weekdays
    for(d in 1:length(weekday)){
      # Read data file for purpose/weekday
      raw_data <- read_delim(file = glue("TTS16-data-inputs/OD-GGH4-by-Time-Day-Mode-Purpose/ttsCross-OD-GGH4-{purpose[p]}-{weekday[d]}-{mode[m]}.txt"), 
                             delim = "\t", 
                             col_names = FALSE,
                             show_col_types = FALSE)

@@

        # Bind the data frames
        od_purpose_time <- rbind(od_purpose_time,
                                 od_purpose_temp)
      }else{
        # Print warning for manual check
        print(glue("TTS16-data-inputs/OD-GGH4-by-Time-Day-Mode-Purpose/Table ttsCross-OD-GGH4-{purpose[p]}-{weekday[d]}-{mode[m]}.txt is empty: check"))
        empty_tables <- empty_tables |>
          mutate(status = ifelse(day == weekday[d] & prps == purpose[p] & md == mode[m],
                                 "Not readable", status))
      }
    }  
  }
}

@@

# Loop through tables that were not successfully read using the previous protocol
for(t in 1:nrow(empty_tables)){
  # Read data file for purpose/weekday
  raw_data <- read_delim(file = glue("TTS16-data-inputs/OD-GGH4-by-Time-Day-Mode-Purpose/ttsCross-OD-GGH4-{empty_tables$prps[t]}-{empty_tables$day[t]}-{empty_tables$md[t]}.txt"), 
                         delim = "\t", 
                         col_names = FALSE,
                         show_col_types = FALSE,
                         skip_empty_rows = FALSE)

@@

      od_purpose_temp2 <- read_csv(glue("TTS16-data-inputs/OD-GGH4-by-Time-Day-Mode-Purpose/ttsCross-OD-GGH4-{empty_tables$prps[t]}-{empty_tables$day[t]}-{empty_tables$md[t]}.txt"),
                                   skip = idx[i] + 1,
                                   n_max = (idx[i+1]-3) - (idx[i]+2),
                                   show_col_types = FALSE,
                                   name_repair = "minimal") |>
        rename(ggh4_orig = 1) |>
        pivot_longer(cols = -ggh4_orig,
                     names_to = "ggh4_dest",
                     values_to = "Trips") |>
        mutate(start_time = start_time$start_time[i],
               day = empty_tables$day[t],
               purpose = empty_tables$prps[t],
               mode = empty_tables$md[t])

@@

Bind the tables:
```{r}
od_ggh4_mode_purpose <- rbind(od_purpose_time,
                         od_purpose_temp) 
```

Summary of table:
```{r}
summary(od_ggh4_mode_purpose)
```

Make start times of trips consistent with day. Some trips are recorded on a certain day, say Monday, but with a start time that exceed midnight (i.e., 24:00). Those trips should be recorded as having been made on the following day. This code corrects this quirk of the data:
```{r}
od_ggh4_mode_purpose <- od_ggh4_mode_purpose |>
  # Convert days to numeric; the convention is that Sunday is 1
  mutate(day = case_when(day == "Monday" ~ 2,
                         day == "Tuesday" ~ 3,
                         day == "Wednesday" ~ 4,
                         day == "Thursday" ~ 5,
                         day == "Friday" ~ 6),
         # Check whether the starting time is greater than 24:00; if true, calculate the time as if it was on the following day, otherwise keep as is
         start_time = if_else(start_time > hm("24:00"), 
                              start_time - hm("24:00"), 
                              start_time),
         # Check whether the starting time is greater than 24:00; if true, change the day of the trip to the following day, otherwise keep as is
         day = if_else(start_time >= hm("24:00"), 
                       day + 1, 
                       day),
         # Convert days from numeric to weekdays
         day = wday(day,
                    label = TRUE))
```

Summary of table:
```{r}
summary(od_ggh4_mode_purpose)
```

Some trips are now recorded as having taken place on Saturday. In general, a small number of trips are changed to the following day, but now the maximum start time is 24:00.

Total number of trips in table:
```{r}
sum(od_ggh4_mode_purpose$Trips)
```

There is a small number of trips ($n = 7,368$) that where taken for an unknown purpose by any mode and on any day. There are zero trips that where undertaken by an unkown mode for any purpose on any day.

In other words, the table above is incomplete, but the number of trips that are missing is very small.

Rename columns to make it consistent with other data objects in package:
```{r}
od_ggh4_mode_purpose <- od_ggh4_mode_purpose |>
  rename(Origin = ggh4_orig,
         Destination = ggh4_dest)
```

Save the data:
```{r}
usethis::use_data(od_ggh4_mode_purpose, 
                  overwrite = TRUE)
```

<!---

-->

Load data object:
```{r}
library(TTS2016R)
data("od_ggh4_mode_purpose")
```

@@

Summary of table:
```{r}
summary(od_ggh4_mode_purpose)
```

Check the total number of trips in the survey:
```{r}
sum(od_ggh4_mode_purpose$Trips)
```

Plot frequency by start time of trips:
```{r}
ggplot(data = od_ggh4_mode_purpose |>
         filter(day != "Sat") |>
         mutate(start_time = as.numeric(start_time, "hours")),
       aes(x = start_time,
           weight = Trips,
           color = day)) +
  geom_density(bw = 0.5) +
  ggtitle("Trip start time for all purposes")# +
#facet_wrap(~ purpose)
```

Plot frequency by start time of trips:
```{r}
ggplot(data = od_ggh4_mode_purpose |>
         filter(day != "Sat") |>
         mutate(start_time = as.numeric(start_time, "hours")),
       aes(x = start_time,
           weight = Trips,
           color = day)) +
  geom_freqpoly(bins = 50) +
  ggtitle("Trip start time for all purposes")# +
#facet_wrap(~ purpose)
```

Summary of trips:
```{r}
od_ggh4_mode_purpose |>
  group_by(mode, day) |>
  summarize(Trips = sum(Trips),
            .groups = "drop")
```

```{r}
ggplot(data = od_ggh4_mode_purpose |>
         filter(day != "Sat", purpose == "Home") |>
         mutate(start_time = as.numeric(start_time, "hours")),
       aes(x = start_time,
           weight = Trips,
           color = day)) +
  geom_density(bw = 0.5) +
  ggtitle("Trip start time for home")# +
#facet_wrap(~ purpose)
```

Read geography file:
```{r}
load(file = glue(here::here(), "/Geography Files/gghv5.rda"))
```

@@

Plot trip attraction:
```{r}
ggplot() +
  geom_sf(data = gghv5 |>
            filter(cmaname == "Hamilton"),
          aes(fill = trip_attraction_Monday),
          color = NA) +
  scale_fill_distiller(name = "Trips", direction = 1, trans = "log") +
  ggtitle("Trip attraction for work - Monday") +
  theme_minimal()
```
--->

>>>>

Loop through the tables and check the filters recorded in each:
```{r}
  # d <- 1
# p <- 2
# m <- 6

@@

for(p in 1:length(purpose)){
  # Loop through modes
  for(m in 1:length(mode)){
    # Loop through weekdays
    for(d in 1:length(weekday)){
      # Read the files in turn
      raw_data <- read_delim(file = glue("ttsCross-OD-GGH4-{purpose[p]}-{weekday[d]}-{mode[m]}.txt"), 
                             delim = "\t", 
                             col_names = FALSE,
                             show_col_types = FALSE)

@@

# Loop through purposes
for(p in 1:length(purpose)){
  # Loop through modes
  for(m in 1:length(mode)){
    # Loop through weekdays
    for(d in 1:length(weekday)){
      # Read data file for purpose/weekday
      raw_data <- read_delim(file = glue("ttsCross-OD-GGH4-{purpose[p]}-{weekday[d]}-{mode[m]}.txt"), 
                             delim = "\t", 
                             col_names = FALSE,
                             show_col_types = FALSE)

@@

        # Bind the data frames
        od_purpose_time <- rbind(od_purpose_time,
                                 od_purpose_temp)
      }else{
        # Print warning for manual check
        print(glue("Table ttsCross-OD-GGH4-{purpose[p]}-{weekday[d]}-{mode[m]}.txt is empty: check"))
        empty_tables <- empty_tables |>
          mutate(status = ifelse(day == weekday[d] & prps == purpose[p] & md == mode[m],
                                 "Not readable", status))
      }
    }  
  }
}

@@

# Loop through tables that were not successfully read using the previous protocol
for(t in 1:nrow(empty_tables)){
  # Read data file for purpose/weekday
  raw_data <- read_delim(file = glue("ttsCross-OD-GGH4-{empty_tables$prps[t]}-{empty_tables$day[t]}-{empty_tables$md[t]}.txt"), 
                         delim = "\t", 
                         col_names = FALSE,
                         show_col_types = FALSE,
                         skip_empty_rows = FALSE)

@@

      od_purpose_temp2 <- read_csv(glue("ttsCross-OD-GGH4-{empty_tables$prps[t]}-{empty_tables$day[t]}-{empty_tables$md[t]}.txt"),
                                   skip = idx[i] + 1,
                                   n_max = (idx[i+1]-3) - (idx[i]+2),
                                   show_col_types = FALSE,
                                   name_repair = "minimal") |>
        rename(ggh4_orig = 1) |>
        pivot_longer(cols = -ggh4_orig,
                     names_to = "ggh4_dest",
                     values_to = "Trips") |>
        mutate(start_time = start_time$start_time[i],
               day = empty_tables$day[t],
               purpose = empty_tables$prps[t],
               mode = empty_tables$md[t])

@@

Bind the tables:
```{r}
od_purpose_time <- rbind(od_purpose_time,
                         od_purpose_temp) 
```

@@

Summary of table:
```{r}
summary(od_purpose_time)
```

Check the total number of trips in the survey:
```{r}
sum(od_purpose_time$Trips)
```

Plot frequency by start time of trips:
```{r}
ggplot(data = od_purpose_time |>
         mutate(start_time = as.numeric(start_time, "hours")),
       aes(x = start_time,
           weight = Trips,
           color = day)) +
  geom_density(adjust = 1) +
  ggtitle("Trip start time for all purposes")# +
#facet_wrap(~ purpose)
```

```{r}
ggplot(data = od_purpose_time |>
         filter(purpose == "Home") |>
         mutate(start_time = as.numeric(start_time, "hours")),
       aes(x = start_time,
           weight = Trips,
           color = day)) +
  geom_density(adjust = 2) +
  ggtitle("Trip start time for home")# +
#facet_wrap(~ purpose)
```
Read geography file:
```{r}
load(file = glue(here::here(), "/Geography Files/gghv5.rda"))
```

@@

Plot trip attraction:
```{r}
ggplot() +
  geom_sf(data = gghv5 |>
            filter(cmaname == "Hamilton"),
          aes(fill = trip_attraction_Monday),
          color = NA) +
  scale_fill_distiller(name = "Trips", direction = 1, trans = "log") +
  ggtitle("Trip attraction for work - Monday") +
  theme_minimal()
```
